\phantomsection
\prefacesection{Acknowledgments}

% INSERT YOUR ACKNOWLEDGEMENTS %
%I would like to acknowledge my parents and friends. This thesis, a manifestation of the many sleepless nights, would not have been possible without their support. This section highlights your acknowledgment of individuals whom you believe deserve recognition for their contribution towards making your achievement possible. 

%first draft which more of a placeholder
% Financial
I would like to thank The MasterCard Foundation for funding my four-year undergrad degree, enabling me to achieve my lifelong dream of becoming an engineer. With their financial support and mentorship, I have been able to achive this dream.

% Academic
Words cannot express my gratitude to my supervisor, Dr. Yaaseen, for his invaluable patience and feedback on Machine Learning and Digital Signal Processing techniques for image processing. I also could not have undertaken this journey without my co-supervisor, Dr Lindie, who generously provided knowledge and expertise in Medical Imaging Technologies to enable me to complete this thesis successfully.  

% Friends
I am also grateful to my classmates and cohort members, especially my friends Clifford and Thikazi, for their constant emotional support, late-night feedback sessions, and moral support. Thanks should also go to the biomedical lab technical officer from the university, who guided me closely in learning how to use the Lodox technology. 

% Family
Lastly, I would be remiss in not mentioning my family, especially my parents, Daniel and Violet, and siblings, namely Carlsberg, Bevy and Brasil. Their belief in me has kept my spirits and motivation high during this process. 

\clearpage % keep this for proper page numbering!